Empirical Evaluation of eval in R

### 2020-08-17 inst/analysis/analysis_stable.Rmd
The case of `base`

```{r eval_call_sites_base}
eval_calls %>% filter(caller_package == "base") %>% select(caller_function) %>% n_distinct(na.rm = TRUE)
```

<details>
  <summary> Comparing with <code>grep</code></summary>
  Grepping *base* to find eval call sites yields 40 `eval` call sites.
</details>

```{r eval_call_sites_per_package}
eval_calls %>% filter(caller_package != "base") %>% group_by(caller_package) %>% summarize(nb_call_sites =  n_distinct(eval_call_srcref, na.rm = TRUE)) %>% arrange(desc(nb_call_sites))
```

@@

@@

*base* has the highest number of eval calls, i.e. `r nb_eval_calls_per_package["per_cent", 1]`% of them.

@@

>>>>

To asses how precise the call site estimate is for calls that do not have `srcref`, we compare it on *base* with the actual count given by `grep`.

```{r eval_call_sites_base}
nb_call_site_base <- eval_calls %>% filter(caller_package == "base") %>% select(caller_function) %>% n_distinct(na.rm = TRUE)
```
The estimate for *base* is `r nb_call_site_base` call sites, while `grep` yields 40 `eval` call sites.

For the other packages, ranked by call sites:
```{r eval_call_sites_per_package}
eval_calls %>% filter(caller_package != "base") %>% group_by(caller_package) %>% summarize(nb_call_sites =  n_distinct(eval_call_srcref, na.rm = TRUE)) %>% arrange(desc(nb_call_sites))
```

@@

What are the packages with the highest number of `eval` calls?

@@

*base* has the highest number of eval calls, i.e. `r nb_eval_calls_per_package[1, "per_cent"][[1]]`% of them.

@@

### Excluding *base* from the data set

We tried to replace the 40 `eval` call sites in *base* and could do it in about 50% of the cases. The rest of the cases are legitimate uses of `eval`, basically variations of `eval`.

TODO: show some examples, replacements and non replacements

We argue that `eval` in *base* should be considered apart from the other `eval` calls.

`eval` in *base* does not use `parse`, so does not take strings potentially coming from users as input,  and thus are deemed to be _safe_.

```{r parse_in_base}
nb_parse_base <- eval_calls %>% filter(!is.na(expr_parsed_expression), caller_package == "base")
```
